---
title: "DATA 621 – Business Analytics and Data Mining"
author: "VP"
date: "`r Sys.Date()`"
subtitle: 'Homework 4: Critical Thinking Group 2'
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
## load library
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("class")) install.packages("class") 
if (!require("knitr")) install.packages("knitr")
if (!require("pROC")) install.packages("pROC") 
if (!require("tidyverse")) install.packages("tidyverse") 
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("corrplot")) install.packages("corrplot") 

knitr::opts_chunk$set(echo = TRUE)
options("scipen" = 10)

library(VIM)
library(caret)
library(mice)
```


#### Read Data

Here, we read the training dataset into a dataframe.

```{r}
insurance_tf_train <- read.csv( "https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_train.csv")[-1]
insurance_tf_test <- read.csv("https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_test.csv")[-1]
insurance_tf_eval <- read.csv( "https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_eval.csv")
```


## Linear logistic regression

### Model 1 - All binary variables with TARGET_FLAG. 

```{r m1, echo=FALSE, eval=TRUE}
model1 <- lm(TARGET_FLAG ~.  ,data=insurance_tf_train, family= "binomial")

summary(model1)
```

### Model 2 - All significant variables with TARGET_FLAG. 

```{r m2, echo=FALSE, eval=TRUE}
model2 <- lm(TARGET_FLAG ~ . -AGE -HOMEKIDS -INCOME -EDUCATION..High.School-JOB.Doctor-JOB.Manager-BLUEBOOK-CAR_TYPE.Van-RED_CAR.yes-CAR_AGE ,data=insurance_tf_train, family=binomial())

summary(model2)
```

```{r m3, echo=FALSE, eval=TRUE}
model3 <- lm(TARGET_FLAG ~ . -AGE -HOMEKIDS -INCOME -EDUCATION..High.School-JOB.Doctor-JOB.Manager-BLUEBOOK-CAR_TYPE.Van-RED_CAR.yes-CAR_AGE-SEX.z_F-EDUCATION.PhD-CLM_FREQ-EDUCATION.Masters-JOB.Lawyer ,data=insurance_tf_train, family= "binomial")

summary(model3)
```

## Selection Model

```{r fig1, fig.height = 4, fig.width = 6, fig.align= 'center'}
roc(TARGET_FLAG~model1$fitted.values, data = insurance_tf_train,plot = TRUE, main = "ROC CURVE", col= "blue",
    percent=TRUE,
    ci = TRUE, # compute AUC (of AUC by default)
    print.auc = TRUE)

```
```{r fig2, fig.height = 4, fig.width = 6, fig.align= 'center'}
roc(TARGET_FLAG~model2$fitted.values, data = insurance_tf_train,plot = TRUE, main = "ROC CURVE", col= "blue",
    percent=TRUE,
    ci = TRUE, # compute AUC (of AUC by default)
    print.auc = TRUE)
```

```{r fig3, fig.height = 4, fig.width = 6, fig.align= 'center'}
roc(TARGET_FLAG~model3$fitted.values, data = insurance_tf_train,plot = TRUE, main = "ROC CURVE", col= "blue",
    percent=TRUE,
    ci = TRUE, # compute AUC (of AUC by default)
    print.auc = TRUE)
```



Based the fact that the area under the curve for model 2 and model 3 are virtually identical. I am going to select model2 Because Auc value is little bit higher than other.

